---
title: "Model_explanation"
output: html_document
date: "2025-03-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries

```{r}
library(lme4)      # For multilevel modeling
library(lmerTest)  # For p-values in multilevel models
library(ggplot2)   # For data visualization
library(dplyr)     # For data manipulation
library(sjstats)   # For calculating ICC
library(sjPlot)    # For plotting model results
library(modelsummary)

df_reduced <- readRDS("df_reduced.rds")

# Quick data check
table(df_reduced$country_name)
```

### Visualize qc19 by country

```{r}
# Calculate the percentage of "Yes" (1) responses for qc19 by country
country_means <- aggregate(qc19 == 1 ~ country_name, data = df_reduced, FUN = mean)

# Multiply by 100 to get percentage and the sort
country_means$percentage <- country_means$`qc19 == 1` * 100
country_means <- country_means[order(country_means$percentage), ]

# Create the plot
ggplot(country_means, aes(x = reorder(country_name, percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Support for Transgender Rights by Country",
       subtitle = "Percentage answering 'Yes' to allowing transgender people to change civil documents",
       x = "Country",
       y = "Support (%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


### Prepare for GLM

```{r}
df_reduced <- df_reduced %>%
  mutate(
    age_z = scale(age),
    religion_z = scale(as.numeric(religion)),
    political_ideology_z = scale(political_ideology),
    respondent_cooperation_z = scale(as.numeric(respondent_cooperation)),
    rainbow_score_z = scale(rainbow_score_2019),
    v2x_egaldem_z = scale(v2x_egaldem))

# Recode qc19 from 1/2 to 0/1
df_reduced$qc19 <- ifelse(df_reduced$qc19 == 1, 1, 0)

# Optimize the settings to avoid convergence problems
control = glmerControl(optimizer = "bobyqa", 
                       optCtrl = list(maxfun = 100000))

```


## 1. Null model (intercept only)

There is significant variation in qc19 scores between countries, justifying the need for a multilevel approach. About 12% of the differences in qc19 scores can be attributed to country-level factors, while 88% is due to individual differences. 12% of variation on the country-level is considered a lot, as the threshold for
going ahead with multilevel models is usually 5%.

```{r}
# STEP 1: Null Model (Intercept-only model)
null_model <- glmer(qc19 ~ 1 + (1|country_name), family = "binomial", data = df_reduced)
summary(null_model)

# Calculate ICC
performance::icc(null_model)
```

## 2. Individual-level predictors (level 1)

Individual characteristics and interview quality (respondent cooperation) significantly predict qc19 scores. The strongest predictors are having LGB friends and respondent cooperation. Religion has a negative impact, suggesting that more religious individuals tend to have lower qc19 scores. Even after controlling for these individual factors, significant country-level differences remain.

```{r}
# STEP 2: Add Individual-Level Predictors (Level 1)
# starting with predictors that showed correlation with qc19
model1 <- glmer(qc19 ~ age_z + gender + religion_z + political_ideology_z + lgb_friends + trans_friends + respondent_cooperation_z + (1|country_name), 
               family = "binomial", 
               control = control,
               data = df_reduced)
summary(model1)

# compare with the null_model
modelsummary(
  list("Model 1" = null_model, 
       "Model 2" = model1),
  stars = TRUE,
  gof_map = c("nobs", "aic", "bic", "logLik", "r.squared"))
```

## 3. Country-level predictors (level 2)

Country-level factors, particularly those related to LGBTQ+ rights (rainbow score) and egalitarian democracy, strongly explain why qc19 scores differ between countries. The negative relationships suggest that more progressive and egalitarian countries tend to have lower qc19 scores. These two country-level variables explain a remarkably large portion (74.5%) of the between-country differences.

The change from gender_equality_index to v2x_egaldem has provided a stronger country-level effect, suggesting that broader measures of egalitarianism in a society may be more relevant to explaining variation in qc19 than gender equality specifically.

```{r}
# STEP 3: Add Country-Level Predictors (Level 2)
model2 <- glmer(qc19 ~ age_z + gender + religion_z + political_ideology_z + lgb_friends + trans_friends +       respondent_cooperation_z +
                 rainbow_score_z + v2x_egaldem_z + (1|country_name), 
               family = "binomial", 
               control = control,
               data = df_reduced)
summary(model2)

# compare with the previous two models
modelsummary(
  list("Model 1" = null_model, 
       "Model 2" = model1,
       "Model 3" = model2),
  stars = TRUE,
  gof_map = c("nobs", "aic", "bic", "logLik", "r.squared"))
```

## 4. Random slopes for individual-level predictors

The effect of having LGB friends on qc19 scores significantly varies across countries. This confirms that the relationship between having LGB friends and qc19 is not uniform but depends on the country context. The negative correlation between random intercepts and slopes indicates that in countries with higher baseline qc19 scores, having LGB friends makes less of a difference. In contrast, in countries with lower baseline qc19 scores, having LGB friends has a stronger positive effect.

The significant random slope suggests that we should proceed with examining cross-level interactions to explain why the effect of having LGB friends varies across countries. Both country-level variables (rainbow_score_2019 and v2x_egaldem) remain highly significant in explaining country-level differences in qc19 scores.

```{r}
# STEP 4: Add Random Slopes for Individual-Level Predictors
# Let's add random slopes for lgb_friends which has shown a strong effect
model3 <- glmer(qc19 ~ age_z + gender + religion_z + political_ideology_z + lgb_friends + trans_friends + respondent_cooperation_z + 
                rainbow_score_z + v2x_egaldem_z + 
                (1 + lgb_friends|country_name), 
               family = "binomial", 
               data = df_reduced)
summary(model3)

# compare with previous three models
modelsummary(
  list("Model 1" = null_model, 
       "Model 2" = model1,
       "Model 3" = model2,
       "Model 4" = model3),
  stars = TRUE,
  gof_map = c("nobs", "aic", "bic", "logLik", "r.squared"))
```

## 5. Cross-level interactions

Neither interaction term (lgb_friends with rainbow_score_2019 or with v2x_egaldem) was statistically significant. This means that while the effect of having LGB friends varies across countries (as shown by the significant random slope in Model 3), this variation isn't explained by differences in countries' LGBTQ+ rights or egalitarian democracy levels.

```{r}
# STEP 5: Add Cross-Level Interactions
# Assuming the random slope for lgb_friends is significant
# We'll test cross-level interactions with both country-level predictors
model4a <- glmer(qc19 ~ age_z + gender + religion_z + political_ideology_z + lgb_friends + trans_friends + respondent_cooperation_z + 
                rainbow_score_z + v2x_egaldem_z + 
                lgb_friends:rainbow_score_z +
                (1 + lgb_friends|country_name), 
               family = "binomial", data = df_reduced)
summary(model4a)

# Alternative interaction with v2x_egaldem
model4b <- lmer(qc19 ~ age_z + gender + religion_z + political_ideology_z + lgb_friends + trans_friends + respondent_cooperation_z + 
                rainbow_score_z+ v2x_egaldem_z + 
                lgb_friends:v2x_egaldem_z +
                (1 + lgb_friends|country_name), 
               data = df_reduced)
summary(model4b)

# compare with previous models
modelsummary(
  list("Model 1" = null_model, 
       "Model 2" = model1,
       "Model 3" = model2,
       "Model 4" = model3,
       "Model 5a" = model4a,
       "Model 5b" = model4b),
  stars = TRUE,
  gof_map = c("nobs", "aic", "bic", "logLik", "r.squared"))
```

## 6. Compare models

The model comparison confirms that Model 3 (with random slopes for lgb_friends) is optimal. Each step from the null model through Model 3 shows significant improvement in fit (p \< 0.001), with dramatic improvements when adding individual predictors (Chisq = 2123.55), country-level predictors (Chisq = 39.48), and random slopes (Chisq = 29.74). However, adding cross-level interactions in Models 4a and 4b did not significantly improve model fit (p = 0.586 for the rainbow score interaction). Model 3 also has the lowest BIC (52398), confirming it strikes the best balance between explanatory power and parsimony. This supports our decision to focus detailed analysis on Model 3.

```{r}
#compare all models
anova(null_model, model1, model2, model3, model4a, model4b)
```

## 7. Variance explained

Our multilevel model explains a substantial portion of variance in qc19 scores. It accounts for 72.5% of between-country differences and 99.3% of within-country individual variation, for a total of 96.1% explained variance. The model confirms that both individual factors (particularly having LGB friends) and country-level characteristics (rainbow scores and egalitarian democracy) significantly predict qc19, while also showing that the effect of having LGB friends meaningfully varies across countries.

```{r}
# 1. Calculate proportion of variance explained
# Get variance components from null model
var_null <- as.data.frame(VarCorr(null_model))
between_var_null <- var_null$vcov[1]  # Between-country variance
within_var_null <- var_null$vcov[2]   # Within-country (residual) variance

# Get variance components from final model (model3)
var_model3 <- as.data.frame(VarCorr(model3))
between_var_model3 <- var_model3$vcov[1]  # Between-country variance
within_var_model3 <- var_model3$vcov[2]   # Within-country (residual) variance

# Calculate proportion of variance explained
between_var_explained <- (between_var_null - between_var_model3) / between_var_null
within_var_explained <- (within_var_null - within_var_model3) / within_var_null
total_var_explained <- 1 - ((between_var_model3 + within_var_model3) / (between_var_null + within_var_null))

cat("Proportion of between-country variance explained:", round(between_var_explained * 100, 1), "%\n")
cat("Proportion of within-country variance explained:", round(within_var_explained * 100, 1), "%\n")
cat("Proportion of total variance explained:", round(total_var_explained * 100, 1), "%\n")
```
