---
title: "Imputation Analysis"
author: "Phong, Marvin, Isabel"
date: "2025-03-04"
output: html_document
---

```{r}
library(haven)
library(tidyverse)
library(ggplot2)
library(xml2)
library(stringr)
library(mice)
```

# Missing Data Overview

```{r}
# codebook <- read_html("../data/raw/ESS9e03_2 codebook.html")
# 
# attr_list <- xml_find_all(codebook,".//h3") %>% 
#   xml_contents() %>% 
#   as_list() %>% 
#   unlist() %>% 
#   .[-1:-2]
# 
# attr_desc <- xml_find_all(codebook,".//h3/following-sibling::*[1]") %>% 
#   as_list() %>% 
#   unlist() %>% 
#   .[-1:-2]
# 
# attr_df <- cbind(attr_list, attr_desc) %>% 
#   as_data_frame()
```

```{r}
data <- read_dta("../data/raw/ZA7575.dta")

data_labels <- sapply(data, function(x) attr(x, "label")) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "attribute") %>% 
  rename("description" = 2)

missing_data <- colSums(is.na(data)) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "attribute") %>% 
  rename("na_count" = 2) %>% 
  filter(na_count > 0) %>% 
  arrange(desc(na_count)) %>% 
  left_join(data_labels, join_by("attribute")) %>% 
  mutate(missing_pcg = na_count/nrow(data),
         question = str_extract(attribute, "\\w\\w?\\d{1,2}")) 

missing_data

summary(missing_data)
```

Overall percentage of missing values

```{r}
sum(missing_data$missing_pcg) / ncol(data)
```

Missing values per variable

```{r}
missing_data %>% 
  group_by(question) %>% 
  summarise(avg_missing_pcg = mean(missing_pcg))  %>% 
  arrange(desc(avg_missing_pcg))
```

-   p6 - Size of locality: has the highest missing data on average (96%) almost most are not answering this question. This might not be highly relevant to the analysis

-   p13 - Language of interview: has the second highest amount of missing data (94%). This question may be helpful for the analysis.

-   p7 - Region: high amount of missing data (94%). We might be able to fill in this question from other questions such as postal code.

-   qc3 - Thinking about the most recent time when you felt discriminated against, under what circumstances did it take place?: high amount of missing data (84%) and highly important for the purpose of the analysis

-   d40 - Could you tell me how many people aged 15 years or more live in your household, yourself included?: high amount of missing data (78%) might not be very relevant for analysis

-   qa3 - Why do you think you are not benefitting from international trade? (M): moderate amount of missing data (67%). The question might not be relevant to the analysis.

-   qa15 - From which of the following countries outside of the EU have you bought goods or services? moderate amount of missing data (62%) but might not be relevant to the analysis

-   d15 (specifically d15 b) - Did you do any paid work in the past? What was your last occupation?: moderate missing data (52%) but could be relevant for the analysis

-   qa2 - Why do you think you are benefitting from international trade? (M): low to moderate missing data (37%). The question is also highly correlated to qa3.

-   qb8 - Did the EU energy label have an influence on the choice of your purchase of electric appliances (fridges, washing machines, dish washers, televisions…)?: low amount of missing data (8%)

-   qa5 (specifically b) - Following up qa5a Which of the following statements best describes your opinion about globalisation? Firstly?: low amount of missing data (7%)

-   qa18 (specifically b) - Following up qa18a Where do you get most of your information on globalisation and international trade? Firstly: low amount of missing data (5%)

-   qa4 (specifically b) - Following up qa4a In your opinion, what should be the main priorities of the European Union's trade policy for the years to come? Firstly? (M): low amount of missing data (3%)

--\> keeping only qa5 and qc3

Missing Data Mechanism

# Imputation

## Multiple Imputation with ML

```{r}
df <- data %>% 
  select(!(starts_with("p6")|starts_with("p13")|starts_with("p7")|starts_with("d40")|starts_with("qa3")|starts_with("qa15")|starts_with("d15b")|starts_with("qa2")|starts_with("qb8")|starts_with("qa18b")|starts_with("qa4b")))

var <- sapply(df, var) %>% 
  as.list() %>% 
  .[. != 0 | is.na(.)] %>% 
  names()

df <- df %>% 
  select(all_of(var)) %>% 
  select(!c("doi", "version", "caseid", "uniqid", "serialid")) %>% 
  mutate(across(c(isocntry), as.factor)) %>% 
  sjlabelled::remove_all_labels()
```

```{r}
df_rf <- complete(mice(df, m = 3, method='rf', maxit = 3, seed = 8532), action=3)
```

```{r}
write_rds(df_rf, "df_rf.rds")
```

## Imputed data analysis

```{r}
imputed_cols <- df %>% 
  select(any_of(c(unique(missing_data$attribute))))
```

## KS Test

## Chi-Square Test

# Sensitivity Analysis
