---
title: "Survey II - Final"
author: "Phong, Marvin, Isabel"
date: "2025-03-01"
output: html_document
---

# Libraries

```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
```

# Cross-country Difference Analysis

Objectives:

-   Analyze the data to understand why support for transgender individuals obtaining official documents varies significantly between countries

-   Individual-level indicators complemented with country-level factors (e.g., cultural norms, legal frameworks. historical context, economic variables, societal attitudes towards LGBTI+ rights)

## Cultural Norms Context

```{r}
rb_cat <- read_csv(file = "../data/2024-rainbow-map-data.csv", n_max = 2) %>% 
  select(-1,-2,-3) %>% 
  t(.) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "category") %>% 
  rename("criteria" = 2,
         "ranking_point" = 3) %>% 
  mutate(category = gsub("\\.\\.\\.\\d+", "", category),
         ranking_point = parse_number(gsub(",", ".", ranking_point))) 
  # separate(criteria, into = c("criteria", "group"), sep = "\\(") %>% 
  # mutate(group = gsub("\\)", "", group)) %>% 
  # mutate(group = ifelse(is.na(group), "General", group))

rb_data <- read_csv(file = "../data/2024-rainbow-map-data.csv", skip = 1) %>% 
  rename("country_code" = 1,
         "country" = 2,
         "ranking" = 3) %>% 
  drop_na(country) %>% 
  mutate(ranking = ranking/100)
```

```{r}
happiness <- read_rds("../happiness_scores.rds")

ggplot(happiness, aes(x=reorder(Country, Happiness_Score, decreasing = TRUE), y=Happiness_Score)) +
  geom_col() +
  xlab("Country") +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}

```

# Predictive Model

Objectives:

-   Develop a predictive model using ML techniques to forecast support levels based on observed trends

-   Ensure model's accuracy and reliability, rigorous calibration, and validation.

## Relevant Features

## Modeling Method

## Model Training

## Model Performance Evaluation
